---
title: "Grade Calculator"
author: "Vincent Quenneville-Bélair"
date: "February 5, 2016"
output: html_document
---

We begin by packages to handle opening spreadsheets, to tidy the data, and to manipulate the data frames.
```{r echo=FALSE, message=FALSE}
# install.packages(c("tidyr", "dplyr")
# install.packages(c("xlsx", "readODS"))

library(xlsx)
library(readODS)

library(stringr)
library(tidyr)
library(dplyr)
```

# Loading and cleaning the data

We load the first spreadsheet, and modify the ID command to get a number.
```{r}
labels <- c("ID", "Lab 1", "Lab 2", "Lab 3", "Lab 4", "Lab 5", "Lab 6", "Lab 7", "Lab 8", "Lab 9", "Lab 10", "Lab 11", "Lab 12", "Quiz 2", "Quiz 1")

# Load spreadsheet
sec12 <- read.xlsx("data/Fall 2015 Lab 120, Sec 1L12.xlsx", sheetName = "Sheet1", rowIndex = -c(1:7), colIndex = c(1:15))
# summary(sec12)

# Assign the correct labels
names(sec12) <- labels
sec12 <- sec12[complete.cases(sec12[,1]),]

# Clean IDs
sec12 <- sec12 %>%
  mutate(ID = str_replace_all(ID, "s", "")) %>%
  mutate(ID = str_replace_all(ID, " ", "")) %>%
  mutate(ID = str_replace_all(ID, "S", ""))

sec12$ID <- as.integer(sec12$ID)
```

We now read the second spreadsheet, and do similar conversions.
```{r}
# Load spreadsheet
sec09 <- read.ods("data/Fall 2015 Lab 120 section 1L09.ods", sheet = 1)
sec09 <- sec09[9:28,1:15]
names(sec09) <- labels

# Clean IDs
sec09 <- sec09 %>%
  mutate(ID = str_replace_all(ID, "s", "")) %>%
  mutate(ID = str_replace_all(ID, " ", "")) %>%
  mutate(ID = str_replace_all(ID, "S", ""))

# Convert to numeric data type
sec09 <- sapply(sec09, as.numeric)
sec09 <- as.data.frame(sec09)
sec09$ID <- as.integer(sec09$ID)
```

We finally read the third spreadsheet.
```{r}
# Load spreadsheet
lec <- read.xlsx("data/lecture.xlsx", sheetName = "Sheet1", colIndex = c(2:5))
labels <- c("ID", "Midterm 1", "Midterm 2", "Final 1")
names(lec) <- labels

# Clean IDs
lec <- lec %>%
  mutate(ID = str_replace_all(ID, "s", "")) %>%
  mutate(ID = str_replace_all(ID, " ", "")) %>%
  mutate(ID = str_replace_all(ID, "S", ""))
```

We then merge the two sections, and merge with lecture data.
```{r}
data <- rbind(sec09, sec12)
data <- merge(data, lec, by="ID", all = TRUE)
```

We scale the two quiz columns to 100.
```{r}
data$`Quiz 1` = data$`Quiz 1`/20*100
data$`Quiz 2` = data$`Quiz 2`/20*100

# Backup original dataset
data.original <- data
```

# Tidying the data

We tidy the data, and make the data frame narrow and long.
```{r}
data <- data %>% 
  gather(key = "Type", value = "Score", -ID) %>%
  separate(col = "Type", into = c("Type", "TypeNumber")) 

# Convert to numerics
data$TypeNumber <- as.integer(data$TypeNumber)
# Remove not necessary column
data <- select(data, -`TypeNumber`)

# Backup tidy data
data.tidy = data
```

We now treat all the missing scores as 0.
```{r}
data[is.na(data)] <- 0
```

# Computing the weigthed score

We drop the two lowest labs.
```{r}
data <- data %>%
  group_by(ID, Type) %>%
  mutate(RankedScore = rank(Score, na.last = FALSE, ties.method = "first")) %>%
  filter((RankedScore > 2 & Type == "Lab") | (Type !="Lab")) %>%  # Drop two lowest labs
  select(-`RankedScore`)

data.bestlabs <- data
```

We make the weighted sum of the two midterms. We also take the mean per ID and Type.
```{r}
weighted_minmax <- function(x){
  .20/.55 * min(x) + .35/.55 * max(x)
}

data[data$Type == "Midterm",] <- data[data$Type == "Midterm",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = weighted_minmax(Score))
  
# Take mean per ID and Type
data <- data %>% 
  group_by(ID, Type) %>% 
  summarise_each(funs(mean))
```

We now scale each score to its corresponding weight in the final score.
```{r}
data[data$Type == "Lab",] <- data[data$Type == "Lab",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .8 * .15 * Score)

data[data$Type == "Quiz",] <- data[data$Type == "Quiz",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .2 * .15 * Score)

data[data$Type == "Midterm",] <- data[data$Type == "Midterm",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .55 * .85 * Score)
  
data[data$Type == "Final",] <- data[data$Type == "Final",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .45 * .85 * Score)
```

We can now simply sum per ID to get the final scores.
```{r comment=""}
data %>% 
  select(-Type) %>%
  group_by(ID) %>% 
  summarise_each(funs(sum))
```
