---
title: "Grade Calculator"
author: "Vincent Quenneville-Bélair"
date: "February 5, 2016"
runtime: shiny
output: html_document
---

```{r echo=FALSE, message=FALSE}
library(knitr)
```

`r opts_chunk$set(echo=FALSE, message=FALSE)`

The goal is to compute the final score of each student. We need to open the required spreadsheets, tidy and manipulate the data, and finally compute the weighted score.
```{r}
# install.packages(c("tidyr", "dplyr")
# install.packages(c("xlsx", "readODS"))

library(xlsx)
library(readODS)

library(stringr)
library(tidyr)
library(dplyr)

library(ggplot2)
library(shiny)
```

## Loading and cleaning the data

We load the first spreadsheet, and modify the ID command to get a number.
```{r}
# Load spreadsheet
sec12 <- read.xlsx("data/Fall 2015 Lab 120, Sec 1L12.xlsx", sheetName = "Sheet1", rowIndex = -c(1:7), colIndex = c(1:15))
# summary(sec12)

# Assign the correct labels
names(sec12) <- names(sec12) %>% str_replace("\\.{1,}", " ") %>% str_to_title()
names(sec12)[1] <- "ID"
names(sec12)[7] <- "Lab 6"

# Removing entries without IDs
sec12 <- sec12[complete.cases(sec12[,1]),]

# Clean IDs
sec12 <- sec12 %>%
  mutate(ID = str_replace_all(ID, "s", "")) %>%
  mutate(ID = str_replace_all(ID, " ", "")) %>%
  mutate(ID = str_replace_all(ID, "S", ""))

sec12$ID <- as.integer(sec12$ID)

head(sec12)
#   ID Lab 1 Lab 2 Lab 3 Lab 4 Lab 5 Lab 6 Lab 7 Lab 8 Lab 9 Lab 10 Lab 11 Lab 12 Quiz 2 Quiz 1
# 1 21    70    85    97    90   100    90    80    90    95    100   89.7     NA      6      3
# 2 22    95   100    90    90    90   100    97   100   100    100   96.2     NA     16     15
# 3 23    70    80    90    93    90    90    85    NA    95    100  100.0     NA      6      5
# 4 24    90    70    NA    90    NA    NA    NA    NA    NA     NA     NA     NA     NA     NA
# 5 25    70    90    97    97    95    95    85    NA   100     85  100.0     NA     10     10
# 6 26    NA    NA    NA    NA    NA    NA    NA    NA    NA     NA     NA     NA     NA     NA
```

##

We now read the second spreadsheet, and do similar conversions.
```{r}
# Load spreadsheet
sec09 <- read.ods("data/Fall 2015 Lab 120 section 1L09.ods", sheet = 1)
sec09 <- sec09[9:28,1:15]

# Assign the correct labels
names(sec09) <- names(sec12)

# Clean IDs
sec09 <- sec09 %>%
  mutate(ID = str_replace_all(ID, "s", "")) %>%
  mutate(ID = str_replace_all(ID, " ", "")) %>%
  mutate(ID = str_replace_all(ID, "S", ""))

# Convert to numeric data type
sec09 <- sapply(sec09, as.numeric)
sec09 <- as.data.frame(sec09)
sec09$ID <- as.integer(sec09$ID)

head(sec09)
#   ID Lab 1 Lab 2 Lab 3 Lab 4 Lab 5 Lab 6 Lab 7 Lab 8 Lab 9 Lab 10 Lab 11 Lab 12 Quiz 2 Quiz 1
# 1  1    80    NA    90    NA    NA    70    NA    NA    NA     NA     NA     NA     NA     NA
# 2  2   100   100    95   100   100   100   100   100   100    100     95    100      6    6.0
# 3  3   100   100    97   100   100    79   100   100   100     80     97    100     18   13.0
# 4  4    80    80    93    90    90    78    95   100    NA     NA     93     90     NA    2.0
# 5  5   100   100    93   100   100    86   100   100    95     80     93    100     14    8.5
# 6  6    90    NA    95    95   100    80   100   100    95     60     95     95      3    5.0
```

##

We finally read the third spreadsheet.
```{r}
# Load spreadsheet
lec <- read.xlsx("data/lecture.xlsx", sheetName = "Sheet1", colIndex = c(2:5))
labels <- c("ID", "Midterm 1", "Midterm 2", "Final 1")
names(lec) <- labels

# Clean IDs
lec <- lec %>%
  mutate(ID = str_replace_all(ID, "s", "")) %>%
  mutate(ID = str_replace_all(ID, " ", "")) %>%
  mutate(ID = str_replace_all(ID, "S", ""))

head(lec)
#   ID Midterm 1 Midterm 2 Final 1
# 1  1        44        94      50
# 2  2        56        95      75
# 3  3        46        91      43
# 4  4       100        89      62
# 5  5        82        90      42
# 6  6        55        74      43
```

##

We then merge the two sections, and merge with lecture data.
```{r}
data <- rbind(sec09, sec12)
data <- merge(data, lec, by="ID", all = TRUE)
```

We scale the two quiz columns to 100.
```{r}
data$`Quiz 1` = data$`Quiz 1`/20*100
data$`Quiz 2` = data$`Quiz 2`/20*100

# Backup original dataset
data.original <- data

head(data)
#   ID Lab 1 Lab 2 Lab 3 Lab 4 Lab 5 Lab 6 Lab 7 Lab 8 Lab 9 Lab 10 Lab 11 Lab 12 Quiz 2 Quiz 1 Midterm 1
# 1  1    80    NA    90    NA    NA    70    NA    NA    NA     NA     NA     NA     NA     NA        44
# 2  2   100   100    95   100   100   100   100   100   100    100     95    100     30   30.0        56
# 3  3   100   100    97   100   100    79   100   100   100     80     97    100     90   65.0        46
# 4  4    80    80    93    90    90    78    95   100    NA     NA     93     90     NA   10.0       100
# 5  5   100   100    93   100   100    86   100   100    95     80     93    100     70   42.5        82
# 6  6    90    NA    95    95   100    80   100   100    95     60     95     95     15   25.0        55
#   Midterm 2 Final 1
# 1        94      50
# 2        95      75
# 3        91      43
# 4        89      62
# 5        90      42
# 6        74      43
```

## Tidying the data

We tidy the data, and make the data frame narrow and long.
```{r}
data <- data %>% 
  gather(key = "Type", value = "Score", -ID) %>%
  separate(col = "Type", into = c("Type", "TypeNumber")) 

# Convert to numerics
data$TypeNumber <- as.integer(data$TypeNumber)

# Backup tidy data
data.tidy <- data

# Remove not necessary column
data <- select(data, -`TypeNumber`)

# Look at data
head(data)
#   ID Type Score
# 1  1  Lab    80
# 2  2  Lab   100
# 3  3  Lab   100
# 4  4  Lab    80
# 5  5  Lab   100
# 6  6  Lab    90
```

We now treat all the missing scores in types other than labs as 0.
```{r}
r <- data$Type != "Lab"
c <- "Score"
data[r, c][is.na(data[r, c])] <- 0
```

## Computing the weigthed score

We drop the two lowest labs.
```{r}
data <- data %>%
  group_by(ID, Type) %>%
  mutate(RankedScore = rank(Score, na.last = FALSE, ties.method = "first")) %>%
  filter((RankedScore > 2 & Type == "Lab") | (Type !="Lab")) %>%  # Drop two lowest labs
  select(-`RankedScore`)

data.bestlabs <- data
```
If there are NAs still assigned for labs, the student did not complete enough labs.

We make the weighted sum of the two midterms. We also take the mean per ID and Type.
```{r}
weighted_minmax <- function(x){
  .20/.55 * min(x) + .35/.55 * max(x)
}

data[data$Type == "Midterm",] <- data[data$Type == "Midterm",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = weighted_minmax(Score))
  
# Take mean per ID and Type
data <- data %>% 
  group_by(ID, Type) %>% 
  summarise_each(funs(mean))

head(data)
#      ID    Type     Score
#   (int)   (chr)     (dbl)
# 1     1   Final  50.00000
# 2     1     Lab        NA
# 3     1 Midterm  75.81818
# 4     1    Quiz   0.00000
# 5     2   Final  75.00000
# 6     2     Lab 100.00000
```

We now scale each score to its corresponding weight in the final score.
```{r}
data[data$Type == "Lab",] <- data[data$Type == "Lab",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .8 * .15 * Score)

data[data$Type == "Quiz",] <- data[data$Type == "Quiz",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .2 * .15 * Score)

data[data$Type == "Midterm",] <- data[data$Type == "Midterm",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .55 * .85 * Score)
  
data[data$Type == "Final",] <- data[data$Type == "Final",] %>% 
  group_by(ID, Type) %>% 
  mutate(Score = .45 * .85 * Score)
```

We can now simply sum per ID to get the final scores.
```{r}
data <- data %>% 
  select(-Type) %>%
  group_by(ID) %>% 
  summarise_each(funs(sum))

head(data)
#      ID   Score
#   (int)   (dbl)
# 1     1      NA
# 2     2 79.3700
# 3     3      NA
# 4     4 79.4130
# 5     5 70.2395
# 6     6 59.7525
```

If the score is NA, the student did not do enough lab to pass the course.
```{r}
data[is.na(data)] <- 0

# Backup final course scores
data.course <- data

head(data)
#      ID   Score
#   (int)   (dbl)
# 1     1  0.0000
# 2     2 79.3700
# 3     3  0.0000
# 4     4 79.4130
# 5     5 70.2395
# 6     6 59.7525
```

Merge the final course score in the tidy table.
```{r}
data$Type <- "Course"
data$TypeNumber <- 1
data <- rbind(data.tidy, data)

# Save data frame to CSV
write.csv(data, "course.csv")
```

## Grade distribution

```{r}

column(2, selectInput("TypeHist", "Type:", c(unique(as.character(data$Type))), selected = "Course"))
renderPlot({
    x    <- na.omit(data[data$Type == input$TypeHist, "Score"])
    
    num_bins = 10
    bins <- seq(min(x), max(x), length.out = num_bins + 1)

    # draw the histogram with the specified number of bins
    qplot(x, breaks = bins, col=I("white"))
  })
```

## Interactive gradebook

```{r}
# Make two dropdown menu appear side-by-side
column(2, selectInput("ID", "ID:", c("All", unique(as.character(data$ID)))))

# Display the resulting table
DT::renderDataTable(DT::datatable({
    
    data <- data %>% 
      unite("Type", Type, TypeNumber, sep = " ") %>% 
      spread(Type, Score)
    
    names(data)[2] <- "Course"
    names(data)[3] <- "Final"

    data$X <- NULL
    if (input$ID != "All") {
      data <- data[data$ID == input$ID,]
    }
    data
  }, rownames = FALSE))
```
